---
title: "dhs2db_tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dhs2db_tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Using dhs2db

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Permission and requirements

This package and tutorial assume that you have an account at dhsprogram.com and permission to access datasets.

## Set the environment

First, load the required packages into your R session.

```{r setup}
library(dhs2db)
library(rdhs)
library(sf)
```


You may wish to keep your personal information outside of your codebase, and `.Renviron` is a good way to accomplish this. Each line in the `.Renviron` file will allow you to set an environment variable (for example, mine has a single line `DHS_EMAIL = <your_email@example.com>`). I use an environment email to store, for example, usernames and passwords and include the `.Renviron` in my `.gitignore` file, so that sensitive information remains on my local computer and is never uploaded to the cloud.

You must restart your R session to load environment variables from the `.Renviron` folder. To follow this tutorial without restarting R, you can set environment variables within the R session by running the following code


```{r eval = FALSE}

Sys.setenv(DHS_EMAIL = "your_email@example.com")
Sys.setenv(DHS_PROJECT = "YOUR DHS PROJECT NAME")


```


## Load the data into R

We will load information into R using the `rdhs` package. For detailed information on retrieving data using `rdhs`, visit the [vignette](https://cran.r-project.org/web/packages/rdhs/vignettes/introduction.html) This package assumes you download data in STATA as zip files. When prompted, type `1` That you would write files outside your R temporary directory, then enter the password for your dhsprogram.com account.

```{r warning=FALSE}

# Set up credentials
set_rdhs_config(email = Sys.getenv("DHS_EMAIL"),
                project = Sys.getenv("DHS_PROJECT"),
                config_path = "rdhs.json",
                cache_path = "C:\\Users\\JWROZE~1\\AppData\\Local\\jwrozelle\\rdhs\\Cache2",
                # cache_path = "dhs_mlm_project",
                global = F)


# Get DHS surveys
survs <- dhs_surveys(
  countryIds = "UG",
  surveyType = "DHS",
  surveyStatus = "Completed",
  surveyYear = c(2016)
)

# download the datasets
datasets <- dhs_datasets(
  surveyIds = survs$SurveyId, 
  fileFormat = "stata",
  force = TRUE
  )
downloads <- get_datasets(
  datasets$FileName, 
  clear_cache = TRUE)


# make an empty list of dhs data
dhsData.list <- list()

# load each file into R
for (svyTable in names(downloads)) {
  
  # load the downloaded data (sometimes there are weird permissions issues with rDHS. For the sake of this tutorial we will ignorre them if they fail)
  try(mySvy <- readRDS(downloads[[svyTable]]))
  
  if (exists("mySvy")) {
    # assign the name of the download
    dhsData.list[[svyTable]] <- mySvy
  }
  
  # remove the generic survey variable from the environment
  rm(mySvy)
}

dhsData.list[["UGGE7AFL"]] <- st_read("C:/Users/jwrozelle/Downloads/UGGE7AFL/UGGE7AFL.shp")



```
##



